State of the Practice for Medical Imaging Software Based on Open Source Repositories

### 2024-11-09 StateOfPractice/Papers/SOP_MI_CritReviewsBioMedEng/SOP_MI_OSR.tex
Updates to section 8

We draw out the point points the points experienced by MI
software developers, compare them to those of general research software
developers, and what practices are used to address these pain points.

Interviews identified the following issues:
\begin{enumerate*}
\item lack of time and funding,
\item technology hurdles, 
\item correctness, and 
\item usability.  
\end{enumerate*}
We first detail each pain point, and contrast this experience with that
from other domains. We then separately cover potential way to mitigate these
issues.

@@

\item[P\refstepcounter{pnum}\thepnum \label{P_LackDevTime}:] \textbf{Lack of
Development Time:} This was felt to be the most significant obstacle, and
is not uncommon
\cite{PintoEtAl2018, PintoEtAl2016, WieseEtAl2019}.
%\cite{SmithEtAl2024}

\item[P\refstepcounter{pnum}\thepnum \label{P_LackFunding}:] \textbf{Lack of
Funding:} As research software is not yet seen as essential infrastructure,
there is an ongoing struggle to attract funding and, even harder, maintain
research software. This is a common complaint%
\cite{GewaltigAndCannon2012, Goble2014, KaterbowAndFeulner2018, SmithEtAl2024}.
As software output is not always counted when judging the academic excellence
of applicants, which forces researchers who write software to spend extra
time on publicity~\cite{WieseEtAl2019}. In some domains, like biology%
\cite{HowisonAndBullard2016}, software is infrequently cited even when used
pervasively. In other words, there is a lack of a formal reward system for
research software~\cite{PintoEtAl2018}.

\item[P\refstepcounter{pnum}\thepnum \label{P_TechnologyHurdles}:]
\textbf{Technology Hurdles:} Developers mentioned: hard to keep up with changes
in OS and libraries, difficult to transfer to new technologies, hard to support
multiple OSes, and hard to support lower-end
computers. Developers expressed difficulty balancing between four factors:
cross-platform compatibility, convenience of development and maintenance,
performance, and security.

\item[P\refstepcounter{pnum}\thepnum \label{P_Correctness}:]
\textbf{Correctness:} The most mentioned thread to correctness was complexity.
Example sources of complexity are: Interviewees identified multiple threats to
the large variety of data formats, complicated data standards, differing
outputs between medical imaging machines, and the addition of (non-viewing
related) functionality.  Other identified threats to correctness:

@@

At least to some extent the problems for MI software users are due to holes in
their knowledge of software and computing technology, and sometimes also
lack expertize in their domain to be able to adequately use the software
\cite{WieseEtAl2019}. 
%A similar pattern has been observed during interviews with
%LBM software developers [citation redacted for blind review].
%\cite{SmithEtAl2024}.

@@

\item[P\ref{P_LackFunding}:] \textbf{Lack of Funding:}
One interviewee proposed an interesting idea: Licensing the software
to commercial companies to integrate it into their products.
In general, solutions need to be more systemic, and we are seeing the rise
of ``research software engineers'' in various jurisdictions.

@@

A number of developers perceive that some of the ``shifting technologies''
problems could be alleviated by moving to web-based applications.
Most of the teams (83\%) chose to develop
native applications. Three of the eight we interviewed were
building web applications, and another team was considering it too.

\item[P\ref{P_Correctness}:] \textbf{Correctness:}
Testing was the most often strategy for ensuring correctness, and teams
mentioned test related activities, including test-driven development, component
tests, integration tests, smoke tests, regression tests, self tests and
automated tests.  This puts MI software ahead of other domains where
insufficient testing is a problem~\cite{PintoEtAl2018} and
insufficiently well-understood~\cite{HannayEtAl2009}.

Another correctness strategy mentioned multiple times is a development process
that involves stable releases and nightly builds.  Other strategies that were
mentioned:
\begin{enumerate*}
\item using CI/CD, 
\item using de-identified copies of medical images for
debugging, 
\item sending beta versions to medical workers who can access the data to
do the tests, 
\item collecting/maintaining a dataset of problematic images,
\item using open datasets,
\item if (part of) the team belongs to a medical school or a hospital, using the
datasets they can access,
\item if the team has access to MRI scanners, self-building sample images for
testing, and
\item if the team has connections with MI equipment manufacturers, asking for
their help on data format problems.
\end{enumerate*}

>>>>

Interviews identified the following issues:
\begin{enumerate*}
\item lack of time,
\item lack of funding,
\item technology hurdles, 
\item correctness, and 
\item usability.  
\end{enumerate*}
We first detail each pain point, and contrast this experience with that
from other domains. We then separately cover potential way to mitigate these
issues.

@@

\item[P\refstepcounter{pnum}\thepnum \label{P_LackDevTime}:] \textbf{Lack of
Development Time:} This was felt to be the most significant obstacle, and
is not uncommon
\cite{PintoEtAl2018, PintoEtAl2016, WieseEtAl2019}.
%DOUBLE BLIND \cite{SmithEtAl2024} [redacted for blind review]

\item[P\refstepcounter{pnum}\thepnum \label{P_LackFunding}:] \textbf{Lack of
Funding:} As research software is not yet seen as essential infrastructure,
there is an ongoing struggle to attract funding to build and, even harder, maintain
research software. This is a common complaint~\cite{GewaltigAndCannon2012, Goble2014, KaterbowAndFeulner2018}. %, SmithEtAl2024 %[redacted for double blind review]
Software output is not always counted when judging the academic excellence of
academics, which forces researchers who write software to spend extra time on
publicity~\cite{WieseEtAl2019}. In some domains, like
biology~\cite{HowisonAndBullard2016}, software is infrequently cited even when
used pervasively. In other words, there is a lack of a formal reward system for
research software~\cite{PintoEtAl2018}.

\item[P\refstepcounter{pnum}\thepnum \label{P_TechnologyHurdles}:]
\textbf{Technology Hurdles:} Developers mentioned the following challenges: hard
to keep up with changes in OS and libraries, difficult to transfer to new
technologies, hard to support multiple OSes, and hard to support lower-end
computers. Developers expressed difficulty balancing between four factors:
cross-platform compatibility, convenience of development and maintenance,
performance, and security.

\item[P\refstepcounter{pnum}\thepnum \label{P_Correctness}:]
\textbf{Correctness:} The most mentioned thread to correctness was complexity.
Example sources of complexity include: a large variety of data formats,
complicated data standards, differing outputs between medical imaging machines,
and the addition of (non-viewing related) functionality.  Other identified
threats to correctness:

@@

At least to some extent the problems for MI software users are due to holes in
their knowledge of software and computing technology, and sometimes also
lack of expertize in their domain to be able to adequately use the software
\cite{WieseEtAl2019}. 
%A similar pattern has been observed during interviews with
%LBM software developers [citation redacted for blind review].
%\cite{SmithEtAl2024}.

@@

\item[P\ref{P_LackFunding}:] \textbf{Lack of Funding:}
One interviewee proposed an interesting idea: Licensing the software
to commercial companies to integrate it into their products.
In general, solutions need to be more systemic, thus we are seeing the rise
of ``research software engineers'' in various jurisdictions.

@@

A number of developers perceive that some of the ``shifting technologies''
problems could be alleviated by moving to web-based applications.
Most of the teams (83\%) chose to develop
native applications. Three of the eight we interviewed were
building web applications, and another team was considering it.

\item[P\ref{P_Correctness}:] \textbf{Correctness:}
Testing was the most often mentioned strategy for ensuring correctness.  Teams
mentioned several test-related activities, including test-driven development, component
tests, integration tests, smoke tests, regression tests, self tests and
automated tests.  This puts MI software ahead of other domains where
insufficient testing is a problem~\cite{PintoEtAl2018} and
insufficiently well-understood~\cite{HannayEtAl2009}.

Another correctness strategy mentioned multiple times is a development process
that involves stable releases and nightly builds.  Other strategies that were
mentioned include:
\begin{enumerate*}
\item using CI/CD, 
\item using de-identified copies of medical images for
debugging, 
\item sending beta versions to medical workers who can access the data to
do the tests, 
\item collecting/maintaining a dataset of problematic images,
\item using open datasets,
\item if (part of) the team belongs to a medical school or a hospital, using the
datasets they can access,
\item if the team has access to MRI scanners, self-building sample images for
testing, and
\item if the team has connections with MI equipment manufacturers, asking for
their help on data format problems.
\end{enumerate*}
